Cover letter for the original IJRR submission

### 2017-05-09 body/contact.tex
typos

Contact generation is typically adressed by randomly sampling a limb configuration, before projecting the effector onto the closest surface with an inverse kinematics solver.
This process is repeated until either a solution for Eq.~(\ref{eq:contact_gen}) is found, or the generator fails, according to a user-defined maximum number of trials.
We favor a modified implementation of this naive approach, more computationally efficient, introduced in our previous work~\citep{Tonneau2014}.

>>>>

Contact generation is typically addressed by randomly sampling a limb configuration, before projecting the effector onto the closest surface with an inverse kinematics solver.
This process is repeated until either a solution for Eq.~(\ref{eq:contact_gen}) is found, or the generator fails, according to a user-defined maximum number of trials.
We favor a modified implementation of this naive approach, more computationally efficient, introduced in our previous work~\citep{Tonneau2014}.

### 2017-05-09 body/intro.tex
typos

Once the contact plan is known, efficient approaches exist to generate a dynamically feasible motion~\citep{Carpentier2016}.
 %~ However in general
%~ the contact plan is manually designed by a time consuming trial-and-error procedure. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in cluttered, human-centered environments, which comprise constraining obstacles (stairs...).
In the specific (and simple) case of gaited biped locomotion on flat ground, 
%~ efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. Furthermore, 
choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (the left foot always follows the right foot).
Efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location.
In the general case planning complex contact interactions is extremely challenging:
At any given time a contact must be chosen between infinitely many possibilities (often a combinatorial discrete choice for the effector and contact surface, and a continuous choice for the contact location). Furthermore, a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles; we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. 
%~ Lastly, the contact manifold has a null measure and is thus impossible to sample. This restriction prevents the use of efficient sampling-based planners.
The
foliation prevents the use of efficient sampling-based
planners
for two reasons. (i) First, each submanifold of the         
foliation
has
a zero measure and cannot be directly sampled.         
A
sample
is
rather obtained by sampling a “free flying”         
configuration
and
explicitly projecting it in contact (which is a         
costly
numerical
operation).
(ii) Second, the foliated topology       
turns
the
exploration
by
spreading
a graph of configurations        
(probabilistic
roadmap,
rapidely
exploring
random trees) into      
an
unefficient
random
process,
where
many
useless nodes       
are
sampled
on
parallel
submanifolds.
The
total
algorithmic       
complexity
of
classical
contact
planners
comes
from
both the        
number
of
graph
nodes
sampled
during
exploration
(ii) and
the cost of the projection when sampling new configurations
(i).

@@

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling has been shown by Escande et al. \cite{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$), and then addressed \Pb as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution could then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. Bouyarmane et al. ~\cite{Bouyarmane2009} precisely focused on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected to the contact manifold using an inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key postures in the path. Chung and Khatib~\cite{7140082} also proposed a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with a discretized environment. This planning phase does not account for collisions, implying that replanning is required in case of failure. In highly-constraining cases such as the car egress scenario we address, we believe that including collision constraints in the planning is a requirement~\citep{tonneauisrr15,grey2017footstep}.

>>>>

Once the contact plan is known, efficient approaches exist to generate a dynamically feasible motion~\citep{Carpentier2016}.
 %~ However in general
%~ the contact plan is manually designed by a time consuming trial-and-error procedure. Automatically addressing the acyclic contact planning problem is thus a pre-requisite to the deployment of robots in cluttered, human-centered environments, which comprise constraining obstacles (stairs...).
In the specific (and simple) case of gaited biped locomotion on flat ground, 
%~ efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location. Furthermore, 
choosing the effector with which to create a contact is trivial because walking follows a cyclic pattern (the left foot always follows the right foot).
Efficient tools such as the capture point~\citep{Pratt2006} can be used to compute the next contact location.
In the general case planning complex contact interactions is extremely challenging:
At any given time a contact must be chosen between infinitely many possibilities (often a combinatorial discrete choice for the effector and contact surface, and a continuous choice for the contact location). Furthermore, a contact choice constrains kinematically and dynamically the possible motions, and there is no analytical way to verify whether this choice brings the robot one step closer to the desired goal or to a dead end, especially in the presence of obstacles; we say that the contact manifold is foliated~\citep{simeon-manipulation-04}. 
%~ Lastly, the contact manifold has a null measure and is thus impossible to sample. This restriction prevents the use of efficient sampling-based planners.
The
foliation prevents the use of efficient sampling-based
planners
for two reasons. (i) First, each sub-manifold of the         
foliation
has
a zero measure and cannot be directly sampled.         
A
sample
is
rather obtained by sampling a “free flying”         
configuration
and
explicitly projecting it in contact (which is a         
costly
numerical
operation).
(ii) Second, the foliated topology       
turns
the
exploration
by
spreading
a graph of configurations        
(probabilistic
road-map,
rapidly
exploring
random trees) into      
an
inefficient
random
process,
where
many
useless nodes       
are
sampled
on
parallel
sub-manifolds.
The
total
algorithmic       
complexity
of
classical
contact
planners
comes
from
both the        
number
of
graph
nodes
sampled
during
exploration
(ii) and
the cost of the projection when sampling new configurations
(i).

@@

\textbf{On the other hand, the two problems \Pa and \Pb might be decoupled} to reduce the complexity. The feasibility and interest of the decoupling has been shown by Escande et al. \cite{DBLP:conf/iser/EscandeKMG08} who manually set up a rough root guide path (i.e. an ad-hoc solution to $\mathcal{P}_1$), and then addressed \Pb as the combinatorial computation of a feasible contact sequence in the neighborhood of the guide. A solution could then be found, %efficiently when considering quadruped locomotion \citep{kalakrishnan2011learning}, 
but at the cost of prohibitive computation times (up to several hours) for constraining scenarios. This approach is suboptimal because the quality of the motion depends on the quality of the guide path. Bouyarmane et al. ~\cite{Bouyarmane2009} precisely focused on automatically computing a guide path with guarantees of \textit{equilibrium feasibility}, by extending key frames of the path into whole-body configurations in static equilibrium. Randomly-sampled configurations are projected to the contact manifold using an inverse-kinematics solver, a computationally-expensive process (about 15 minutes to compute a guide path in the examples presented). Moreover this explicit projection is insufficient to guarantee the feasibility between two key postures in the path. Chung and Khatib~\cite{7140082} also proposed a decoupled approach, with a planning phase based on the reachable workspace of the robot limbs, used to judge the ability to make contact with a discretized environment. This planning phase does not account for collisions, implying that re-planning is required in case of failure. In highly-constraining cases such as the car egress scenario we address, we believe that including collision constraints in the planning is a requirement~\citep{tonneauisrr15,grey2017footstep}.

### 2017-05-09 body/overview.tex
typos

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed sequentially: From a given problem (a) we first plan a root guide path (b), before
extending it into a sequence of static equilibrium configurations(c). In the case where step c fails,
our framework invalidates the computed guide path, and restarts the planning from b.

@@

\subsection{Computation of a guide path --- \Pa (Section~\ref{rbprm})}
We first consider the problem of planning a root guide path (Figure~\ref{fig:framework}--$\mathcal{P}_1$).  The dimension of the path is equal to the number of degrees of freedom (DoFs) of the root of the robot.
Similarly to previous work~\citep{Bouyarmane2009} the path must be \textit{equilibrium feasible}: there must exist a joint configuration that results in static equilibrium for each root configuration\footnote{Enforcing static equilibrium is a classical, conservative approach to reduce the search space and considering states  with non-zero accelerations and velocities, which can't be connected trivially. This does not mean that the final motion will necessary be quasi-static.}. Previous works verifiy \textit{equilibrium feasibility} by explicitly computing such a configuration. We preserve the low dimensionality of the problem by approximating \textit{equilibrium feasibility} with \textit{contact rechability}, illustrated in the following.

>>>>

Figure~\ref{fig:framework} illustrates our work flow.
\Pa and \Pb are addressed sequentially: From a given problem (a) we first plan a root guide path (b), before
extending it into a sequence of static equilibrium configurations(c). In the case where step c fails,
our framework invalidates the computed guide path, and restarts the planning from b.

@@

\subsection{Computation of a guide path --- \Pa (Section~\ref{rbprm})}
We first consider the problem of planning a root guide path (Figure~\ref{fig:framework}--$\mathcal{P}_1$).  The dimension of the path is equal to the number of degrees of freedom (DoFs) of the root of the robot.
Similarly to previous work~\citep{Bouyarmane2009} the path must be \textit{equilibrium feasible}: there must exist a joint configuration that results in static equilibrium for each root configuration\footnote{Enforcing static equilibrium is a classical, conservative approach to reduce the search space and considering states  with non-zero accelerations and velocities, which can't be connected trivially. This does not mean that the final motion will necessary be quasi-static.}. Previous works verify \textit{equilibrium feasibility} by explicitly computing such a configuration. We preserve the low dimensionality of the problem by approximating \textit{equilibrium feasibility} with \textit{contact reachability}, illustrated in the following.
